---
title: "Lab 5"
author: "Sophia Leiker"
date: "10/26/2021"
output: 
  html_document: 
    toc: yes
    theme: cerulean
    number_sections: yes
    code_folding: hide
---

```{r setup, include=FALSE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

# Confidence intervals

In this section we will learn to use the 't.test()' function as a way to return the confidence interval using the t-distribution.

## Example 1

Create a mock sample of observation values, and find the confidence interval. Then we will learn how to report that

```{r}
pika_mass <- c(30, 40, 50, 70, 86, 99, 320, 30, 40)

pika_mean <- mean(pika_mass)

pika_sd <- sd(pika_mass)

pika_ci <- t.test(pika_mass)

pika_ci
```

### Risky and tedius way: 
Mean pika mass at Paiute Pass is 91.9 g $\pm$ 13.5 g (mean $\pm$ 1 standard deviation, n = 7) with a 95% confidence interval of [79.4, 104.4] g. 

### A high initial investment, but safer & better way

In-line code referencing to stored objects!

Use a single backtick on either side of a lowercase r to create an in-lin bit of code

Mean pika mass at Paiute Pass `r round(pika_mean, 1)` g $\pm$ `r round(pika_sd, 1)` g (mean $\pm$ 1 standard deviation, n = `r length(pika_mass)`) with a 95% confidence interval of [`r round(pika_ci$conf.int[1], 1)`, [`r pika_ci$conf.int[2]`] g. 

Note: In order to get specific elements from an output, you can pull out each individual parameters in the console using the dollar sign such a pika_ci{dollarsign}p.value, or pika_ci{dollarsign}conf.int, or pika_mass, pika_mass[2]

# Two-sample t-test

Use a two-sample t-test to test the null hypothesis that samples were drawn from populations with the same mean (difference in means = 0).

- H0: Means difference = 0
- HA: Means difference is NOT = 0

This is a two-sided t-test because no directionality is implied. 

## Example 1

```{r}
oak_trees <- c(29, 19, 22, 30, 35, 16, 20, 7)
  
pine_trees <- c(48, 32, 41, 40, 56, 70, 30, 19 , 20)
```

Is there a significant difference in mean heights for oak and pine trees? 

```{r}
trees_t <- t.test(oak_trees, pine_trees)

```

The p-value of `r trees_t$p.value` means that there is `r ((trees_t$p.value) * 100)` % chance of finding sample means *at least as different as those I found* by random chance if they were drawn from populations with the same mean height. 

Using the 'mpg' dataset to compare city gas milage for SUVs and compact cars.
#to check the different groups that are within a column --> unique(mpg$class)

```{r}
# Create a subset called compact_suv that only contains observations where the class is suc OR compact

compact_suv <- mpg %>% 
  filter(class %in% c("suv", "compact"))

#to check that this worked, run in the console "unique(compact_suv$class)" to check that there are only two unique values being spit out 
```

```{r}
ggplot(data = compact_suv, aes(x= cty)) +
  geom_histogram(bins = 12) + 
  facet_wrap(~class)

ggplot(data = compact_suv, aes(sample = cty)) +
  geom_qq() + 
  facet_wrap(~class)

#If the QQ plot is a nice linear line, you can say that the observations are fairly normally distributed 
```

```{r}
car_stats <- compact_suv %>% 
  group_by(class) %>% 
  summarize(mean_city = mean(cty), #mean city is the column name, then the second argument is telling the summarize table what summary state to calculate
            sd_city = round(sd(cty), 1) , #you can round using this function
            sample_size = n())

car_stats


```

```{r}
compact_sample <- compact_suv %>% 
  filter(class == "compact") %>% 
  pull(cty)

suv_sample <- compact_suv %>% 
  filter(class == "suv") %>% 
  pull(cty)

cars_t <- t.test(compact_sample, suv_sample)

```

Stats Speak: reject the null hypothesis of equal mean city gas milage

Mean gas milage for compact cars and suvs differs significantly (t(`r cars_t$parameter`) = `r cars_t$statistic`, P < 0.001, $\alpha$ = 0.05).
